---
title: "Assignment-2 Solving LP Model Using R"
author: "Villyal Vidushi"
date: "2024-09-21"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

QUESTION- 1

The Weight Corporation has three branch plants with excess production capacity. Fortunately,
the corporation has a new product ready to begin production, and all three plants have this
capability, so some of the excess capacity can be used in this way. This product can be made in
three sizes--large, medium, and small--that yield a net unit profit of $420, $360, and $300,
respectively. Plants 1, 2, and 3 have the excess capacity to produce 750, 900, and 450 units per
day of this product, respectively, regardless of the size or combination of sizes involved.
The amount of available in-process storage space also imposes a limitation on the production
rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 square feet,
respectively, of in-process storage space available for a day's production of this product. Each
unit of the large, medium, and small sizes produced per day requires 20, 15, and 12 square feet,
respectively. Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, medium, and small sizes, respectively, would be sold per day. At each plant, some employees will need to be laid off unless most of the plant’s excess production capacity can be used to produce the new product. To avoid layoffs if possible, management has decided that the plants should use the same percentage of their excess capacity to produce the new product. Management wishes to know how much of each of the sizes should be produced by each of the plants to maximize profit.




```{r}
# Data matrix variable names

data = matrix(c("a_l", "a_m", "a_s", "b_l", "b_m", "b_s", "c_l", "c_m", "c_s"), ncol = 3, byrow = TRUE)
```

# Interpretation

Constructing 3 x 3 matrix by giving character strings into the matrix row by row in terms of the product sizes and plant labels.

```{r}
# Specify the column names and row names of matrix
colnames(data) = c('Large', 'Medium', 'Small')
rownames(data) = c("Plant 1", "Plant 2", "Plant 3")
```

# Interpretation

Creating  the columns ("Large", "Medium", "Small") and rows ("Plant 1", "Plant 2", "Plant 3") of the matrix.


```{r}
#Assigning to table
final = as.table(data)
```

# Interpretation

Converting the assigned matrix into table format to properly represent tabular data.

```{r}
#Display the table
final
```
# Interpretation

The final table which presents a matrix with corresponding row and column numbers.

#Objective:
Maximize profit for the production of three product sizes (large, medium, small) across three plants. The profits per unit are:

Large:  $420
Medium: $360
Small:  $300

lets Assume That

Products produced by Plant 1 are: 
$$al, am, as$$

Products produced by Plant 2 are: 
$$bl, bm, bs$$

Products produced by Plant 3 are: 
$$cl, cm, cs$$

Let take “p” as production capacity percentage.

The above products are in order of Large, Medium and Small

(a)The decision variables are   
$$al, am, as, bl, bm, bs, cl, cm, and cs$$

(b)The linear programming model 
$$Maximize Z = (420al+ 360am + 300as) + (420bl + 360bm + 300bs) + (420cl + 360cm + 300cs)$$


#Constraints -

1.) Storage Space Forecast Constraint :

Plant 1: $$20al + 15am + 12as ≤ 13000$$
Plant 2: $$20bl + 15bm + 12bs ≤ 12000$$
Plant 3: $$20cl + 15cm + 12cs ≤ 5000$$



2.) Production Capacity Constraints :

Plant 1: $$al + am + as ≤ 750$$
Plant 2: $$bl + bm + bs ≤ 900$$
Plant 3: $$cl + cm + cs ≤ 450$$



3.) Sales Constraints :

Large   : $$al + bl + cl ≤ 900$$
Medium  : $$am + bm + cm ≤ 1200$$
Small.  : $$as + bs + cs ≤ 750$$



4.) Equal Capacity Utilization Constraint :
Every plant needs to utilize the same proportion of its total output. let P be the percentage of capacity used at each plant


$$P = \frac{al + am + as}{750} = \frac{bl + bm + bs}{900} = \frac{cl + cm + cs}{450}$$



5.) Non- Negativity Constraints : 
$$
al \geq 0, \; am \geq 0, \; as \geq 0, \; bl \geq 0, \; bm \geq 0, \; bs \geq 0, \; cl \geq 0, \; cm \geq 0, \; cs \geq 0
$$




Solving the LP Problem using ‘lpSolve’ function 


Step 1: Install and load the ‘lpSolve’ Package.


```{r}
#install.packages("lpSolve")
library ('lpSolve')
```

Step 2: Prepare a vector of the objective function

#Objective function vector variables
```{r}
f.obj = c(420, 360, 300, 420, 360, 300, 420, 360, 300)
```

step 3:

#Constraint matrix with variables
```{r}
f.con = matrix(c(20, 15, 12, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 20, 15, 12, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 20, 15, 12,
                 1, 1, 1, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 1, 1, 1, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 1, 1, 1,
                 1, 0, 0, 1, 0, 0, 1, 0, 0,
                 0, 1, 0, 0, 1, 0, 0, 1, 0,
                 0, 0, 1, 0, 0, 1, 0, 0, 1,
                 900, 900, 900, -750, -750, -750, 0, 0, 0,
                 0, 0, 0, 450, 450, 450, -900, -900, -900,
                 450, 450, 450, 0, 0, 0, -750, -750, -750
                 ), nrow = 12, ncol = 9, byrow = TRUE)


f.con

```

Step 4:

#Creating Constraint directions vector
```{r}
f.dir <- c("<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=")
```

Step 5: 

#vector for constraints
```{r}
f.rhs <- c(13000, 12000, 5000, 750, 900, 450, 900, 1200, 750, 0, 0, 0)
```

```{r}
lp("max", f.obj, f.con, f.dir, f.rhs)

```

#Displaying the solution
```{r}
lp('max', f.obj, f.con, f.dir, f.rhs)$solution
```

